---
title: "`lubridate`: time to live code!"
author: "jean prado for RStudio Education"
date: "Dec 16, 2022"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    css: "style.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load `lubridate`

```{r lubridate}
# load lubridate
```

## The dataset

```{r data}
# load nycflights13
```

### Tables

- **airlines**: *carrier* and *name* information for airlines
- **airports**: airports' metadata with location and timezone information
- **flights**: information for all flights that departed NYC in 2013
- **planes**: metadata for all plane tailnumbers in the FAA aircraft registry
- **weather**: hourly meterological data for LGA, JFK and EWR

```{r tables}
# weather

# flights
```

## Parse date and time

### Selecting relevant columns

```{r parse_1}
# without time_hour


```

### Creating day, month, year

```{r parse_2}

```

### Using hour

```{r parse_3}

```

### Check with column

```{r parse_4}
# with time_hour


```

### Using flights

```{r parse_5}
# full date


# only hour

```

## Extract components

### Basic components

```{r extract_1}
# flights, only date-time column

```

### Other cool stuff

```{r extract_2}
# wday, label, week, quarter

```
